Rule Synthesis for Rule-based Intrusion Detectors

### 2020-05-18 paper/paper.tex
pass

Open-source \nids{}, like Snort~\cite{snort} and
Suricata~\cite{suricata}, provide rulesets for many existing
attacks. Unfortunately, attackers actively look for new mechanisms to
\begin{wrapfigure}[8]{r}{0.21\textwidth}
    \vspace{-2ex}  
    \centering
    \scalebox{0.9}{  
      \begin{tikzpicture}
        \begin{axis}[
            width = 0.3\textwidth,
            height = 0.2\textwidth,            
            bar width=4pt,
            scaled ticks=false,
            ymajorticks=false,
            ybar stacked,
            enlargelimits=0.09,
            clip=true,
            legend style={at={(0.5,-0.15)},anchor=north,legend columns=-1},          
            xtick=data,
            xticklabel style={rotate=45, font=\footnotesize},
            tick pos=left,
            nodes near coords,
            every node near coord/.append style={font=\footnotesize},
            nodes near coords align={vertical},
            symbolic x coords={Jan, Feb, Mar, Apr, May, Jun, Jul, Aug, Sep, Oct, Nov, Dec},]
            \addplot coordinates {(Jan, 114) (Feb, 155) (Mar, 117) (Apr, 159) (May, 110) (Jun, 231) (Jul, 122) (Aug, 167) (Sep, 107) (Oct, 225) (Nov, 148) (Dec, 127)};
        \end{axis}
      \end{tikzpicture}
    }
    \vspace{-5ex}
    \caption{\label{fig:distribution-rules-per-month}2019 new rules.}
\end{wrapfigure}
exploit systems through the
network. Figure~\ref{fig:distribution-rules-per-month} shows the
number of rules added per month in the year of 2019 to the ``Emerging
Threats Open Ruleset''~\cite{emerging-threats-open}, a popular public
data set of \suri\ rules. A closer observation on the
changelogs of this ruleset~\cite{emerging-threats-changelogs} shows that new rules are
often created.  Rulesets need to be continuously revised as the
community finds and advertises vulnerabilities, which eventually can be exploited by hackers. 
Unfortunately, creating these rules is
time-consuming and error-prone \rui{is there a citation we could add here?}. The language for describing rules is
non-trivial (see Section~\ref{sec:example-suricata-rules}) and testing
correctness of a rule requires data and careful inspection.

@@

\sloppy This paper proposes \tname{} (for \textbf{Sy}nthesis of
Su\textbf{ri}cata R\textbf{u}le\textbf{s}), a technique to
automatically synthesize rules for rule-based \nids. Our goal is to
facilitate the creation of rules and, consequently, help system
administrators to harden the protection against network attacks.
\tname{} leverages different sources of information to synthesize
rules for a given attack. It uses 1) positive examples, \ie{},
scenarios that reproduce the attack, 2) negative examples in the form
of a data set of benign messages, and 3) a data set of human-written
trusted rules. The positive and negative examples constrain the search
space of rules that can be used to capture the attack. Intuitively,
positive examples indicate what should be captured whereas negative
examples indicate what should not be captured. It is worth noting that
vast amounts of negative data
exist~\cite{tcpreplay,stratosphere-normal}, but positive data is
scarce as it is not common (and considered unethical~\Mar{look for
  citation for ethics code on this}) to publish attacks on the
web\Mar{is there a term for non-deep web?}. Classical machine learning techniques \emph{are not directly
  applicable in this context} for two main reasons: 1)~they require a
large amount of training data to build models (recall that positive
data is scarce) and 2)~they typically have limited support to string
and structured data, which are prevalent in this domain.

@@

We evaluated \tname\ against \Mar{@Lucas, please look for constant
  numbers as this *in text* and use macros instead. We do not want to
  leave inconsistent data in the paper!}14 attacks of various kinds, such as \Fix{...}\rui{add few examples}. The observed results are encouraging. For 
example, \tname\ ranked the correct rule among the top five rules in 
\percTopFiveRanking\ of the cases and ranked the correct rule among 
the top ten rules in 11 of the 14 cases. \rui{to discuss, I find 
this a bit weak.}

@@

>>>>

Open-source \nids{}, like Snort~\cite{snort} and
Suricata~\cite{suricata}, provide rulesets for many existing
attacks. Unfortunately, attackers actively look for new mechanisms to
\begin{wrapfigure}[8]{r}{0.21\textwidth}
    \vspace{-2ex}  
    \centering
    \scalebox{0.9}{  
      \begin{tikzpicture}
        \begin{axis}[
            width = 0.3\textwidth,
            height = 0.2\textwidth,            
            bar width=4pt,
            scaled ticks=false,
            ymajorticks=false,
            ybar stacked,
            enlargelimits=0.09,
            clip=true,
            legend style={at={(0.5,-0.15)},anchor=north,legend columns=-1},          
            xtick=data,
            xticklabel style={rotate=45, font=\footnotesize},
            tick pos=left,
            nodes near coords,
            every node near coord/.append style={font=\footnotesize},
            nodes near coords align={vertical},
            symbolic x coords={Jan, Feb, Mar, Apr, May, Jun, Jul, Aug, Sep, Oct, Nov, Dec},]
            \addplot coordinates {(Jan, 114) (Feb, 155) (Mar, 117) (Apr, 159) (May, 110) (Jun, 231) (Jul, 122) (Aug, 167) (Sep, 107) (Oct, 225) (Nov, 148) (Dec, 127)};
        \end{axis}
      \end{tikzpicture}
    }
    \vspace{-5ex}
    \caption{\label{fig:distribution-rules-per-month}2019 new rules.}
\end{wrapfigure}
exploit systems through the
network. Figure~\ref{fig:distribution-rules-per-month} shows the
number of rules added per month in the year of 2019 to the ``Emerging
Threats Open Ruleset''~\cite{emerging-threats-open}, a popular public
data set of \suri\ rules. A closer observation on the changelogs of
this ruleset~\cite{emerging-threats-changelogs} shows that new rules
are often created.  Rulesets need to be continuously revised as the
community finds and advertises vulnerabilities, which eventually can
be exploited by hackers.  Unfortunately, creating these rules is
time-consuming and error-prone~\cite{Ring_2019}. The language for
describing rules is non-trivial (see
Section~\ref{sec:example-suricata-rules}) and testing correctness of a
rule requires data and careful inspection.

@@

\sloppy This paper proposes \tname{} (for \textbf{Sy}nthesis of
Su\textbf{ri}cata R\textbf{u}le\textbf{s}), a technique to
automatically synthesize rules for rule-based \nids. Our goal is to
facilitate the creation of rules and, consequently, help system
administrators to harden the protection against network attacks.
\tname{} leverages different sources of information to synthesize
rules for a given attack. It uses 1) positive examples, \ie{},
scenarios that reproduce the attack, 2) negative examples in the form
of a data set of benign messages, and 3) a data set of human-written
trusted rules. The positive and negative examples constrain the search
space of rules that can be used to capture the attack. Intuitively,
positive examples indicate what should be captured whereas negative
examples indicate what should not be captured. It is worth noting that
classical machine learning techniques \emph{are not directly
  applicable in this context} as they typically have limited support
to string and structured data, which are prevalent in this domain.

%% vast amounts of negative data
%% exist~\cite{tcpreplay,stratosphere-normal}, but positive data is
%% scarce~\cite{} as it is not common (and considered unethical) to publish
%% attacks on the web\Mar{is there a term for non-deep web?}.

@@

We evaluated \tname\ against \Mar{@Lucas, please look for constant
  numbers as this *in text* and use macros instead. We do not want to
  leave inconsistent data in the paper!}14 attacks of various kinds, such as \Fix{...}\rui{add few examples}. The observed results are encouraging. For 
example, \tname\ ranked the correct rule among the top five rules in 
\percTopFiveRanking\ of the cases and ranked the correct rule among 
the top ten rules in 11 of the 14 cases. \rui{to discuss, I find 
this a bit weak.}\Mar{agree. let us try to add more examples and data.}

@@

\Mar{mention something on ethical hacking.. ethical-hacking}
